---
title: "Downstream Analysis of AnVILHMMRATAC Data"
author:
- name: Kayla Interdonato
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
        toc_flow: true
vignette: >
  %\VignetteIndexEntry{B_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette will walk a user through how to run a full downstream analysis of
HMMRATAC on the output data from the `AnVILHMMRATAC` workflow. The output data
will be retrieved and utilized in functions from the `r Biocpkg("ChIPseeker")`
package.

# Installation

```{r setup}
library(AnVILHMMRATAC)
```

```{r install}
pkgs = c("Bioconductor/AnVIL", "ChIPseeker", "org.Hs.eg.db")
BiocManager::install(pkgs)

suppressPackageStartupMessages({
    library(AnVIL)
    library(ChIPseeker)
    library(org.Hs.eg.db)
})
```

# Obtaining the data

The files that are needed for the `ChIPseeker` analysis are the `.gappedPeak`
files for each sample. We localize those files for the anlysis.

```{r paths}
avworkspace("bioconductor-rpci-anvil/HMMRATAC")
gP_paths <- avworkflow_jobs() %>%
    head(1) %>%
    avworkflow_files() %>%
    filter(endsWith(.data$file, ".gappedPeak")) %>%
    pull(path)

```

`avworkflow_jobs()` will locate the list of jobs run on the AnVIL workspace.
The last job is grabbed using `head(1)`, then the files associated with that
job are listed using `avworkflow_files()`. We know that our sample files all
have the ending `.gappedPeak` so we use `filter(endsWith(.data$file,
".gappedPeak"))` to filter out the files that end with ".gappedPeak". Then we
use `pull(path)` to pull out the paths where the output data was saved on the
Google bucket for this workspace.

With these paths, we then localize them to our local machine to use in the
following analysis.

```{r paths}
localize(gP_paths, getwd(), dry = FALSE)

files_path <- paste0(getwd(), "/SRR8912%.filteredPeaks.gappedPeak")
files <- sprintf(files_path, 69:74)
```

# Downstream analysis

Now we will utilize the `ChIPseeker` package to perform statistical testing of
significant overlap among ChIP seq data sets. It incorporates open access
database GEO to allow users to compare their datasets to those available on
GEO.

```{r chipseeker}
peak <- readPeakFile(file_path)

## display peak locations across entire genome
covplot(peak, weightCol = "weights")

## perform peak annotation
peakAnno <- annotatePeak(files[[4]],
    tssRegion = c(-3000, 3000),
    TxDb = txdb,
    annoDb = "org.Hs.eg.db")

plotAnnoPie(peakAnno)

upsetplot(peakAnno, vennpie = TRUE)

plotDistToTSS(peakAnno,
    title = "Distribution of transcription factor-binding loci relative to
TSS")

# Session Information

```{r}
sessionInfo()
```
